\documentclass[floatfix]{article}  
\pagestyle{plain}

\usepackage{amsmath,amssymb}
\usepackage{dsfont}
\usepackage{graphicx} %loads the graphicx.sty package 
\usepackage{epstopdf} %loads the epstopdf.sty package 
\usepackage{slashed}
\usepackage{color}

\usepackage{graphicx,longtable,tocloft,color}
\usepackage{sidecap}
\usepackage{subfig}
\usepackage{xspace}
\usepackage{mydefs}
\usepackage{cite}
\usepackage{lineno}
\usepackage{multirow}
\usepackage{hyperref}
%\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
%\tolerance = 1000
%\parindent 0cm
%\parskip 2mm
%
%
%for draft only:
%\linenumbers
\usepackage{array}

\title{Constraining models of new physics with collider measurements of Standard Model signatures}

\author{Jonathan M. Butterworth$^1$, David Grellscheid$^2$,\\ Michael Kr\"amer$^3$, David Yallup$^1$\\
\it $^1$Department of Physics and Astronomy, University College London,\\ \it Gower St., London, WC1E 6BT, UK\\
\it $^2$IPPP, Department of Physics,\\\it Durham University, DH1 3LE, United Kingdom\\ \it $^3$Institute for Theoretical Particle Physics and Cosmology, \\ \it RWTH Aachen University, Aachen, Germany}


\begin{document}

\maketitle 

\begin{abstract}
Many searches for striking signatures of physics beyond the Standard Model have been performed at the 
Large Hadron Collider, and to date no compelling evidence for such physics has emerged. 
In addition, large number of precise and detailed
measurements of various final states have been carried out, and the results are generally  
described well by Standard Model calculations.
Using, as examples, simplified models for Dark Matter and R-Parity violating supersymmetry, 
we demonstrate that this library of existing precision measurements can be used to set limits on a wide variety of 
beyond-the-Standard Model physics scenarios. Specifically, we set new limits on these example models using
the inclusive jet and $W+$jet measurements made by ATLAS and CMS in the first run of the LHC 
(maybe add some numbers here on the new limits). Using a wider array
of measurements, including those to come from 13~TeV collisions, is a promising method for exploring 
physics beyond the electroweak symmetry-breaking scale and looking for subtle effects of new physics.
\end{abstract}


\section{Introduction}
\label{sec:intro}

With the discovery of the Higgs boson~\cite{Aad:2012tfa,Chatrchyan:2012ufa}, 
the first data-taking period of the experiments at the 
Large Hadron Collider demonstrated that the understanding of electroweak symmetry-breaking within
the Standard Model is broadly correct, and thus that the theory is potentially valid well above the
TeV scale. Many precision measurements of jets, of vector-boson production, and other of final states, 
have been published, reaching into this new kinematic domain. The predictions of the Standard Model are 
generally in agreement with the data, while the many dedicated searches for physics beyond the Standard Model
have excluded a wide range of possible scenarios. Nevertheless, there are many reasons to be confident that
physics beyond the Standard Model (BSM) exists; examples include the existence of Dark Matter, the large 
preponderance of matter over antimatter in the universe, and the existence of gravity. None of 
these can be easily accommodated within known Standard Model phenomenology. 

This motivates a continued campaign to make precise measurements and calculations at higher energies and 
luminosities, and to exploit these measurements to narrow down the class of viable models of new physics, 
hopefully shedding light on the correct new theory, or at least on the energy scale at which
new physics might be observed at future experiments.

In this paper we exploit three important developments to survey existing measurements and set 
limits on new physics. 

\begin{itemize}
\item
Standard Model predictions for differential and exclusive, or semi-exclusive, 
final states are made using sophisticated calculational software, often embedded in Monte Carlo generators 
capable of simulating full, realistic final states~\cite{Buckley:2011ms}.
These generators now incorporate higher-order processes
matched to logarithmic parton showers and successful models of soft physics such as hadronisation and
underlying event. They are also capable of importing new physics models into this framework, thus allowing
rapid predictions of their impact on a wide variety of final states simultaneously. 
In this paper we make extensive use of these capabilities within Herwig~\cite{Bahr:2008pv}.
\item
As many favoured BSM scenarios have been excluded, there has been a move toward 
``simplified models'' of new physics~\cite{Alves:2011wf}, which aim to be as generic as possible, while being
internally consistent. The philosophy is similar to an ``effective lagrangian'' approach in which effective
couplings are introduced to describe new physics, but is more powerful as the such simplified models also
include new particles, and thus can remain useful up to and beyond the scale of new physics - a region 
potentially probed by LHC measurements.
\item
The precision measurements from the LHC have mostly been made in a manner which minimises the model 
dependence. That is, they are defined in terms of final-state signatures in fiducial regions well matched to the
acceptance of the detector. The majority of such measurements are readily available for analysis and
comparison in the Rivet library~\cite{Buckley:2010ar}. 
\end{itemize}

These three developments together make it possible to efficiently 
bring the power of a very wide range of data to bear on the search for new physics. We focus here on 
some specific examples, which are interesting in themselves and serve to demonstrate the power of the method.

The outline of the paper is as follows. In the next section, we motivate and describe the two 
simplified models to be considered, and their implementation via Herwig. In Section~\ref{sec:measurements}
we introduce the measurements that we will use, and their implementation in Rivet. Section~\ref{sec:kinematics}
discusses the differential cross sections in which the impact of these models would be 
most apparent. In Section~\ref{sec:limits} this impact is translated into limits on the model parameters.

\section{Models}\label{sec:models}


Dark Matter model:
\cite{Kahlhoefer:2015bea}


\section{Measurements}\label{sec:measurements}

\section{Limits}\label{sec:limits}

\section{Conclusions}\label{sec:conclusions}

\section*{Acknowledgments}



\bibliographystyle{h-physrev4}
\bibliography{simple}

\end{document}


